Modeling Predator Preferences

### 2014-09-15 paper/methods.tex
now 2(ST) parameter estimators; lambda, gamma indepdent and l = g*c_st

We denote the binary response that the predator did in fact eat at least one prey species $s$ in time period $t$ by $Z_{jst} = 1(X_{jst} > 0)$.  Now, the observed data $Z_{jst} \iid \text{Bern}(1-\exp\{-\lambda_{st}\})$.  Using this we can find the complete data likelihood, by first noting that the conditional distribution of the unobserved data $X_{jst}$ given $Z_{jst}, Y_{ist}, \boldsymbol{\lambda}, \boldsymbol{\gamma}$ is a truncated Poisson distribution

\begin{equation*}
  f_{X|Y,Z,\boldsymbol{\lambda},\boldsymbol{\gamma}}(x_{jst}) =
  \frac{\exp{\{-\lambda_{st}\}} \lambda_{st}^{x_{jst}}}{(1 - \exp{\{-\lambda_{st}\}}) x_{jst}!}1(x_{jst} \geq 0) \quad \text{ where } \quad \E_{[X|Y,Z]}X_{jst} = \frac{\lambda_{st} \exp{\{\lambda_{st} \}}}{\exp{\{ \lambda_{st} \}} - 1}.
\end{equation*}

@@

\begin{align*}
  \E l_{comp} 
  & = \E \log{f_{X,Z|\boldsymbol{\lambda}}(X_{jst},z_{jst})} + \log{f_{Y|\boldsymbol{\gamma}}(y_{ist})} \\
  & = \sum_{s=1}^S \sum_{t=1}^T \sum_{j=1}^{J_t} \E \log{f_{X,Z|\boldsymbol{\lambda}}(X_{jst},z_{jst})}
  + \sum_{s=1}^S \sum_{t=1}^T \sum_{i=1}^{I_t} \log{f_{Y|\boldsymbol{\gamma}}(y)} \\
  & = \sum_{s,t,j} \left( - \lambda_{st} 
    + z_{jst} \log(\lambda_{st}) \E X_{jst} \right) + \sum_{s,t} \left( -I_t \gamma_{st} + Y_{\cdot st} \log{I_t \gamma_{st}} \right) + \text{const} \\
  & = \sum_{s,t} \left( -J_t \lambda_{st} + z_{\cdot st} \log(\lambda_{st}) \E X_{jst} \right) + \sum_{s,t} \left( -I_t \gamma_{st} + Y_{\cdot st} \log{I_t \gamma_{st}} \right) + \text{const}.
\end{align*}

@@

\noindent where $\theta_0, \theta_1$ represent the parameters estimated under the null and alternative hypotheses, respectively.  It is well known that the asymptotic distribution of $\Lambda$ is a $\chi_{\rho}^2$ distribution with $\rho$ degrees of freedom.  Under the EM algorithm we use $L_{observed}(Z,Y)$ as the likelihood in the calculation of $\Lambda$.

>>>>

We denote the binary response that the predator did in fact eat at least one prey species $s$ in time period $t$ by $Z_{jst} = 1(X_{jst} > 0)$.  Now, the observed data are $Z_{jst} \iid \text{Bern}(1-\exp\{-\lambda_{st}\})$.  Using this we can find the complete data likelihood by first noting that the conditional distribution of the unobserved data $X_{jst}$ given $Z_{jst}, Y_{ist}, \boldsymbol{\lambda}, \boldsymbol{\gamma}$ is a truncated Poisson distribution

\begin{equation*}
  f_{X|Y,Z,\boldsymbol{\lambda},\boldsymbol{\gamma}}(x_{jst}) =
  \frac{\exp{\{-\lambda_{st}\}} \lambda_{st}^{x_{jst}}}{(1 - \exp{\{-\lambda_{st}\}}) x_{jst}!}1(x_{jst} > 0) \quad \text{ where } \quad \E_{[X|Y,Z]}X_{jst} = \frac{\lambda_{st} \exp{\{\lambda_{st} \}}}{\exp{\{ \lambda_{st} \}} - 1}.
\end{equation*}

@@

\begin{align*}
  \E l_{comp} 
  & = \E \log{f_{X,Z|\boldsymbol{\lambda}}(X_{jst},z_{jst})} + \log{f_{Y|\boldsymbol{\gamma}}(y_{ist})} \\
  & = \sum_{s=1}^S \sum_{t=1}^T \sum_{j=1}^{J_t} \E \log{f_{X,Z|\boldsymbol{\lambda}}(X_{jst},z_{jst})}
  + \sum_{s=1}^S \sum_{t=1}^T \sum_{i=1}^{I_t} \log{f_{Y|\boldsymbol{\gamma}}(y)} \\
  & = \sum_{s,t,j} \left( - \lambda_{st} 
    + z_{jst} \E X_{jst} \log{\lambda_{st}}  \right) + \sum_{s,t} \left( -I_t \gamma_{st} + Y_{\cdot st} \log{I_t \gamma_{st}} \right) + \text{const} \\
  & = \sum_{s,t} \left( -J_t \lambda_{st} + z_{\cdot st} \E X_{jst} \log{\lambda_{st}} \right) + \sum_{s,t} \left( -I_t \gamma_{st} + Y_{\cdot st} \log{I_t \gamma_{st}} \right) + \text{const}.
\end{align*}

@@

\noindent where $\theta_0, \theta_1$ represent the parameters estimated under the null and alternative hypotheses, respectively.  It is well known that the asymptotic distribution of $\Lambda$ is a $\chi_{\rho}^2$ distribution with $\rho$ degrees of freedom.  Under the EM algorithm we use $L_{obs}(Z,Y)$ as the likelihood in the calculation of $\Lambda$.
